---
title: "Untitled"
author: "Marc Kissel"
date: "5/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
DCI <- read_csv("DCI_data_full.csv")



library(WDI)



```


```{r}
WDIsearch("*armed") %>%
  as.data.frame() %>%
  tbl_df() %>%
  arrange(str_length(name)) %>%
  View()
```

look at WDI
indicator NY.GDP.PCAP.CD = GDP per capita (current US$)
looks like only data rom 1960 onward for the GDP stuff
might be loosing coutnired in the join due to names...triple check afterwards


```{r}
wdi_indicator <-WDI(indicator = "NY.GDP.PCAP.CD", extra = TRUE, start= 2010, end = 2010)

#i want to rename so i can join. rename DCI

DCI_edited <- DCI %>% mutate(iso3c = stateabb) %>% filter(year == "2010")

joined <- DCI_edited %>% inner_join(wdi_indicator, by = "iso3c") 

#maybe left?

joined %>% count(country)

joined %>%  arrange(desc(NY.GDP.PCAP.CD)) %>% ggplot(aes(x=NY.GDP.PCAP.CD, y = X)) + geom_point()
```

all data

MS.MIL.XPND.GD.ZS = Military expenditure (% of GDP) 
MS.MIL.TOTL.P1	Armed forces personnel, total
MS.MIL.TOTL.TF.ZS	Armed forces personnel (% of total labor force


```{r}
wdi_indicator_all <-WDI(indicator = c( "NY.GDP.PCAP.CD", "MS.MIL.XPND.GD.ZS","MS.MIL.TOTL.TF.ZS" , "VC.PKP.TOTL.UN"), extra = TRUE, start= 1960, end = 2010)

#i want to rename so i can join. rename DCI

DCI_edited_all <- DCI %>% mutate(iso3c = stateabb) %>% filter(year > "1959")

joined_all <- DCI_edited_all %>% inner_join(wdi_indicator_all, by = c("iso3c", "year") )

#i need to check if this work via year?

joined_all %>% filter(year >2000) %>%   arrange(desc(NY.GDP.PCAP.CD)) %>% ggplot(aes(x=NY.GDP.PCAP.CD, y = X)) + geom_point() + facet_wrap(~year) + scale_x_log10() #+ geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE)

#all data

joined_all  %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")

joined_all  %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")


#for the 2000s
joined_all %>% filter(year >2000) %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated") #facet_wrap

joined_all %>% filter(year >2000) %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 2001-2010") # all data post 2000 in one



#1990s

joined_all %>% filter(year >1989 & year <2000) %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated") #facet_wrap


joined_all %>% filter(year >1989 & year <2000) %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "1990-1999 data") #all data


#just one year
joined_all %>% filter(year == 2000) %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x=MS.MIL.XPND.GD.ZS, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "year = 2000") + geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE, angle = 45)

#####################################################
#####for MS.MIL.TOTL.TF.ZS
#####################################################
joined_all  %>%   arrange(desc(MS.MIL.TOTL.TF.ZS)) %>% ggplot(aes(x=MS.MIL.TOTL.TF.ZS, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Armed forces personnel (% of total labor force)
", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")




#for the 2000s
joined_all %>% filter(year >2000) %>%   arrange(desc(MS.MIL.TOTL.TF.ZS)) %>% ggplot(aes(x=MS.MIL.TOTL.TF.ZS, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Armed forces personnel (% of total labor force)
", y = "DCI", title = "DCI and Military expenditure are correlated") #facet_wrap

joined_all %>% filter(year >2000) %>%   arrange(desc(MS.MIL.TOTL.TF.ZS)) %>% ggplot(aes(x=MS.MIL.TOTL.TF.ZS, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Armed forces personnel (% of total labor force)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 2001-2010") # all data post 2000 in one






joined_all %>% filter(year >2000) %>%   arrange(desc(VC.PKP.TOTL.UN)) %>% ggplot(aes(x=VC.PKP.TOTL.UN, y = X)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Presence of peace keepers (number of troops, police...", y = "DCI", title = "DCI and Military expenditure are correlated")




```


#look at model

```{r}
  lm(X  ~ log10(MS.MIL.XPND.GD.ZS),
     data = joined_all) %>%
  summary()

fit1 <-  lm(X  ~ (MS.MIL.XPND.GD.ZS),
     data = joined_all) 

fit2 <-  lm(X  ~ log10(MS.MIL.XPND.GD.ZS),
     data = joined_all) 
```

fit2 (which logs the milt expenditure) shows a sig relationship with milt expen on DCI. as we increase Milt.Exped by 1 unit it increases DCI by .1

Coefficients:
                         Estimate Std. Error t value Pr(>|t|)    
(Intercept)              0.775290   0.001110  698.63   <2e-16 ***
log10(MS.MIL.XPND.GD.ZS) 0.100171   0.002138   46.84   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 0.04556 on 3296 degrees of freedom
  (816 observations deleted due to missingness)
Multiple R-squared:  0.3997,	Adjusted R-squared:  0.3995 
F-statistic:  2194 on 1 and 3296 DF,  p-value: < 2.2e-16



issues: 
some countirs have multiple data in the joined dataframe for a year./

next steps
1. check by decade or see what happens during war times
2. look at other indicatoes


```{r}

joined_all %>% arrange(desc(NY.GDP.PCAP.CD)) %>% ggplot(aes(x=NY.GDP.PCAP.CD, y = X)) + geom_point() + facet_wrap(~year) + scale_x_log10()
```

