---
title: "Untitled"
author: "Marc Kissel"
date: "5/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
DCI <- read_csv("DCI_data_full.csv")



library(WDI)



```


```{r}
WDIsearch("*armed") %>%
  as.data.frame() %>%
  tbl_df() %>%
  arrange(str_length(name)) %>%
  View()
```

look at WDI
indicator NY.GDP.PCAP.CD = GDP per capita (current US$)
looks like only data rom 1960 onward for the GDP stuff
might be loosing coutnired in the join due to names...triple check afterwards


```{r}
wdi_indicator <-WDI(indicator = "NY.GDP.PCAP.CD", extra = TRUE, start= 2010, end = 2010)

#i want to rename so i can join. rename DCI

DCI_edited <- DCI %>% mutate(iso3c = stateabb) %>% filter(year == "2010")

joined <- DCI_edited %>% inner_join(wdi_indicator, by = "iso3c") 

#maybe left?

joined %>% count(country)

joined %>%  arrange(desc(NY.GDP.PCAP.CD)) %>% ggplot(aes(x=NY.GDP.PCAP.CD, y = X)) + geom_point()
```

all data

MS.MIL.XPND.GD.ZS = Military expenditure (% of GDP) 
MS.MIL.TOTL.P1	Armed forces personnel, total
MS.MIL.TOTL.TF.ZS	Armed forces personnel (% of total labor force


```{r}
wdi_indicator_all <-WDI(indicator = c( "NY.GDP.PCAP.CD", "MS.MIL.XPND.GD.ZS","MS.MIL.TOTL.TF.ZS" , "VC.PKP.TOTL.UN"), extra = TRUE, start= 1960, end = 2010)


wdi_indicator_all <- wdi_indicator_all %>% 
  rename(GDP_per_capita = NY.GDP.PCAP.CD,
         Military_expenditure_GDP = MS.MIL.XPND.GD.ZS,
         Armed_forces_personnel=MS.MIL.TOTL.TF.ZS,
         Peacekeepers=VC.PKP.TOTL.UN)

#i want to rename so i can join. rename DCI

DCI_edited_all <- DCI %>% mutate(iso3c = stateabb) %>% filter(year > "1959") #year > 1959 since that is what i can get in WDI

joined_all <- DCI_edited_all %>% inner_join(wdi_indicator_all, by = c("iso3c", "year") )

#i need to check if this work via year?
#UPDATE: what is missing?


joined_all %>% filter(year >2000) %>%   arrange(desc(GDP_per_capita)) %>% ggplot(aes(x=GDP_per_capita, y = X)) + geom_point() + facet_wrap(~year) + scale_x_log10() #+ geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE)

#all data

joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")

joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + geom_smooth(aes(group=1), method =lm)+ scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")

#based on above graph the DCI is higher than expected for many sub-shartan countires based on the model taht precits DCI based on milt expendature.  BUT THAT IS ONLY FOR A CERTAIN YEAR? 
#interesting to think about why that might be. 

p <-joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region, text=Country)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")
library(plotly)
ggplotly(p, tooltip = "text")

#for the 2000s
joined_all %>% filter(year >2000) %>%   arrange(desc(MS.MIL.XPND.GD.ZS)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated") #facet_wrap

joined_all %>% filter(year >2000) %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 2001-2010") # all data post 2000 in one

#####
#what if i facet by region for the 2000s/all
####

joined_all %>% filter(year >2000) %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + facet_wrap(~region) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated")

joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + facet_wrap(~region) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated")





#1990s

joined_all %>% filter(year >1989 & year <2000) %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated") #facet_wrap


joined_all %>% filter(year >1989 & year <2000) %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "1990-1999 data") #all data


#just one year
joined_all %>% filter(year == 2000) %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "year = 2000") + geom_text(aes(label = country), vjust = 1, hjust = 1, check_overlap = TRUE, angle = 45)

#####################################################
#####for MS.MIL.TOTL.TF.ZS
#####################################################
joined_all  %>%   arrange(desc(Armed_forces_personnel)) %>% ggplot(aes(x=Armed_forces_personnel, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Armed forces personnel (% of total labor force)
", y = "DCI", title = "DCI and Armed forces personnel are correlated", subtitle = "from years 1960-2010")




#for the 2000s
joined_all %>% filter(year >2000) %>%   arrange(desc(Armed_forces_personnel)) %>% ggplot(aes(x=Armed_forces_personnel, y = X, color = region)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Armed forces personnel (% of total labor force)
", y = "DCI", title = "DCI and Armed forces personnel are correlated") #facet_wrap

joined_all %>% filter(year >2000) %>%   arrange(desc(Armed_forces_personnel)) %>% ggplot(aes(x=Armed_forces_personnel, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Armed forces personnel (% of total labor force)", y = "DCI", title = "DCI and Armed forces personnel are correlated", subtitle = "from years 2001-2010") # all data post 2000 in one






joined_all %>% filter(year >2000) %>%   arrange(desc(Peacekeepers)) %>% ggplot(aes(x=Peacekeepers, y = X)) + geom_point() + facet_wrap(~year) + scale_x_log10() + labs (x = "Presence of peace keepers (number of troops, police...", y = "DCI", title = "DCI and Military expenditure are correlated")

##############################################
#rayshader
###############


x_graph <- joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X, color = region)) + geom_point() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")

plot_gg(x_graph, height=3, width=3.5, multicore=TRUE, pointcontract = 0.7, soliddepth=-200)

x_graph2 <- joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X)) + geom_hex() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")


plot_gg(x_graph2, height=3, width=3.5, multicore=TRUE, pointcontract = 0.7, soliddepth=-200)



p <-joined_all  %>%   arrange(desc( Military_expenditure_GDP)) %>% ggplot(aes(x= Military_expenditure_GDP, y = X)) + geom_hex() +  scale_x_log10() + labs (x = "Military expenditure (% of GDP)", y = "DCI", title = "DCI and Military expenditure are correlated", subtitle = "from years 1960-2010")

```


#look at model

```{r}
  lm(X  ~ log10(Military_expenditure_GDP),
     data = joined_all) %>%
  summary()

fit1 <-  lm(X  ~ (Military_expenditure_GDP),
     data = joined_all) 

fit2 <-  lm(X  ~ log10(Military_expenditure_GDP),
     data = joined_all) 

fit3 <-  lm(X  ~ log10(Military_expenditure_GDP) + region,
     data = joined_all) 

```

fit2 (which logs the milt expenditure) shows a sig relationship with milt expen on DCI. as we increase Milt.Exped by 1 unit it increases DCI by .1

Coefficients:
                         Estimate Std. Error t value Pr(>|t|)    
(Intercept)              0.775290   0.001110  698.63   <2e-16 ***
log10(MS.MIL.XPND.GD.ZS) 0.100171   0.002138   46.84   <2e-16 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 0.04556 on 3296 degrees of freedom
  (816 observations deleted due to missingness)
Multiple R-squared:  0.3997,	Adjusted R-squared:  0.3995 
F-statistic:  2194 on 1 and 3296 DF,  p-value: < 2.2e-16





```{r}
#broom
library(broom)
tidy(fit2)
```


issues: 
some countirs have multiple data in the joined dataframe for a year./

next steps
1. check by decade or see what happens during war times
2. look at other indicatoes


```{r}

joined_all %>% arrange(desc(NY.GDP.PCAP.CD)) %>% ggplot(aes(x=NY.GDP.PCAP.CD, y = X)) + geom_point() + facet_wrap(~year) + scale_x_log10()
```

##########################
############################

What if i did a map that showed the DCI for every country for a year and over a number of years


UPDATE I THINK MISSING MAP DO TO HOW I FILTER?
```{r}
library(ggthemes)
glimpse(map_data("world"))
joined_all$Country
#are they going to match
map_set <- map_data("world") %>% 
  filter(region != "Antarctica")
DCI_map <- joined_all %>% left_join(map_set, by =(c("Country" = "region")))
#losing the US...

map_set2 <- map_set %>% mutate(region = replace(region, region == "USA", "United States of America"))

DCI_map <- joined_all %>% inner_join(map_set2, by =(c("Country" = "region")))

DCI_map %>% filter(year == 2000) %>% 
  ggplot(aes(x=long, y =lat, group=group, fill = X)) + geom_polygon() +theme_map() + scale_fill_gradient2(low = "blue", mid = "pink", high="red", midpoint = .8)

DCI_map2 <- joined_all %>% inner_join(map_set2, by =(c("Country" = "region")))

DCI_map  %>% 
  ggplot(aes(x=long, y =lat, group=group, fill = X)) + geom_polygon() +theme_map() + scale_fill_gradient2(low = "blue", mid = "pink", high="red", midpoint = .8) 
####################################
#######################################
#6/27 updates:
#i think i joined wrong back in the joined_all step
#go back and check
#########3#
######

joined_all_2 <- DCI_edited_all %>% left_join(wdi_indicator_all, by = c("iso3c", "year") )

DCI_map2 <- joined_all_2 %>% right_join(map_set2, by =(c("Country" = "region")))

DCI_map2 %>% filter(year == 2000) %>% 
  ggplot(aes(x=long, y =lat, group=group, fill = X)) + geom_polygon() +theme_map() + scale_fill_gradient2(low = "blue", mid = "pink", high="red", midpoint = .8)
#####MUCH BETTER
#but it is huge.....

DCI_map2 %>% filter(year == 2000) %>% distinct(Country, year, group, lat, long,  .keep_all = TRUE) %>% ggplot(aes(x=long, y =lat, group=group, fill = X)) + geom_polygon() +theme_map() + scale_fill_gradient2(low = "blue", mid = "pink", high="red", midpoint = .8)


library(gganimate)
DCI_map2 %>% filter(year > 2005) %>% distinct(Country, year, group, lat, long,  .keep_all = TRUE) %>% 
  ggplot(aes(x=long, y =lat, group=group, fill = X)) + geom_polygon() +theme_map() + scale_fill_gradient2(low = "blue", mid = "pink", high="red", midpoint = .8) + transition_time(year) + labs(title = 'Year: {frame_time}')
```

